---
title: "The data model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The data model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## The application database

`ALCHEMER_DB` points at a directory holding two things:

```
$ALCHEMER_DB/
  catalog.sqlite       # the catalog: table structure, snapshots, row locations
  data/                # Parquet data files, written and compacted by DuckDB
```

It's a [DuckLake](https://ducklake.select/) database: Parquet for the data, a
small SQL database tracking table structure, snapshots, and where every row
lives. The catalog is SQLite rather than a DuckDB file because that is what
allows more than one process to use it — analysts can query the database while
the pipeline writes to it, with no lock to wait for. Connect with plain
`DBI::dbConnect()` (see `vignette("getting-started")`) and it attaches
under the alias `alchemer`, with three schemas:

| Schema | What it's for | Written by |
|---|---|---|
| `raw` | The archive. Verbatim, untyped, never edited. | `ingest()` |
| `pub` | Tidy typed tables for analysis. Rebuildable. | `pub_layer()` |
| `meta` | Operational records: runs, decisions, loads, assertions. | both |

In ETL terms `ingest()` is Extract + Load and `pub_layer()` is Transform — an
**ELT** pipeline, because landing the data untouched first is what stops a
surprise from Alchemer breaking a run. In medallion terms, `raw` is bronze and
`pub` is silver.

## `raw`: archival, verbatim, untyped

Every column in `raw` is `VARCHAR` or `JSON`. No coercion, renaming, or
language selection happens at ingest. Alchemer's own typing is inconsistent
enough — `is_test_data` as the string `"1"`/`"0"`, titles as
`{"English": "..."}` maps, a statistics key literally named `"total responses"`
— that doing any of it at ingest would let a new question type fail the whole
run, or silently mangle data. Every table also carries a `payload` column
holding the source object close to verbatim, so nothing this package doesn't
model explicitly is ever lost.

| Table | Grain |
|---|---|
| `raw.surveys` | one row per survey |
| `raw.survey_definitions` | one per survey — the whole definition tree, verbatim |
| `raw.survey_pages`, `raw.survey_questions`, `raw.survey_question_options` | flattened from that tree |
| `raw.responses` | one row per response; answers verbatim in `survey_data` |
| `raw.survey_statistics`, `raw.survey_campaigns` | one row per question / campaign |

### What's never stored

- **File-upload attachment bytes.** The answer value and Alchemer filename are
  kept; the bytes are not downloaded and the expiring download URL is not
  chased. Fetch them yourself if you need them preserved.
- **Respondent IP address, latitude/longitude, city, region, postal code, DMA,
  and user agent** — redacted at ingest, including inside `payload`.
- **Contacts.** Not fetched at all: heavy PII (names, emails, phone numbers)
  that most uses of this package don't need.

## History: snapshots, not a hand-rolled archive

There's no separate versioned layer — DuckLake versions every write. Query the
database as it stood at a past time or version:

```{r}
# con from vignette("getting-started"), which runs USE alchemer -- so raw.*
# and pub.* need no catalog prefix here, though ducklake_snapshots() still
# takes the catalog name as a plain string argument, not a table reference
DBI::dbGetQuery(con, "
  SELECT * FROM raw.responses AT (TIMESTAMP => '2026-01-01 00:00:00')")
DBI::dbGetQuery(con, "SELECT * FROM ducklake_snapshots('alchemer')") # every commit
```

**History depth and retention are the same knob.** `expire_history(db,
older_than)` trims how far back time travel reaches and reclaims the files;
querying an expired version then fails cleanly rather than returning wrong
data. `compact()` merges the small files each refresh leaves behind without
touching history. Schedule both — see `vignette("scheduling")`.

`expunge()` goes further: it removes rows and their history for data that must
actually leave the database rather than just become hard to find. That takes
more than a `DELETE`, because DuckLake keeps deleted rows in their Parquet file
behind a delete marker, inlines small writes into the catalog, and stores
verbatim per-column minimum and maximum values as statistics. `?expunge`
documents all four steps; the short version is that `expunge()` handles them
and a hand-written `DELETE` does not.

## Deletions are flagged, never dropped

Alchemer's list endpoints silently omit deleted responses — there is no
tombstone to detect a deletion from. So every refresh downloads **all** of a
survey's responses and compares that set to what's stored: anything previously
stored and now missing is flagged `is_deleted = TRUE` with a
`deleted_detected_at`, never removed. The same applies to a survey that
disappears from account discovery. A row that reappears upstream is un-flagged
automatically. Filter deleted rows in or out as your analysis needs.

### Only what changed is written

Comparing the fetched set against the stored one is also what keeps the archive
small. DuckLake never rewrites a Parquet file in place, so replacing a survey's
rows wholesale on each refresh would write a fresh copy of every row every
time — for a 5,000-response survey polled every 15 minutes, about 200 MB a day,
all of it retained until snapshots expire. Instead each refresh writes only rows
that actually differ, so an unchanged survey costs nothing but the requests.

One consequence worth knowing: `ingested_at` on a `raw` row means *when this
version of the row was first stored*, not when it was last seen. An untouched
response keeps its original value indefinitely. `meta.runs` and
`meta.run_events` are where "when did we last look" lives.

## Why full refresh, not incremental

`ingest()` always downloads **every** response for a survey it refreshes, never
a filtered subset. An incremental delta keyed on `date_updated` was considered
and rejected: that field carries no timezone suffix, newly submitted data can
take up to five minutes to appear, identical requests are cached for 60
seconds, and deleted responses are omitted regardless. Any one of those can
make a delta miss a row permanently, with nothing downstream able to notice.

Under full refresh, every one of those degrades to "this survey refreshes later
than ideal", and self-heals on the next run. `date_updated`, the survey list's
`modified_on`, and a response count are still used — but only as hints for
**when** a survey is worth refreshing, never to decide **which rows** to keep. A
wrong hint costs freshness, bounded by `full_sweep_days`; it can't cost
correctness.

## `pub`: typed, for ordinary analysis

`pub_layer()` builds typed, language-resolved tables from `raw` alone. It's
rebuildable, incremental by survey, and safe to re-run.

A survey is rebuilt only when something it is built from or with has changed:
its `raw` rows, or the `language`/`tz`/`wide_views` setting, or the version of
alchemeR doing the building. A run where nothing changed upstream does no
per-survey work at all, which is what makes it cheap to call on every pipeline
tick — so `pub_layer()` belongs after every `ingest()`, not on a slower
schedule of its own. Surveys that *are* rebuilt are rebuilt whole, one
transaction each. `pub_layer(force = TRUE)` overrides the skipping.

| Table | Grain | Notes |
|---|---|---|
| `pub.surveys`, `pub.questions`, `pub.options` | one row per entity | typed, language-resolved titles |
| `pub.responses` | one row per response | typed timestamps and booleans; test and deleted rows flagged, not dropped |
| `pub.answers` | one row per response × question | question and option labels, and reporting values, denormalised in |
| `pub.wide_<slug>_<survey_id>` | one row per respondent | a generated view, one per survey |

The wide view's name comes from the survey title (lowercased, ASCII, length
capped) plus the survey id, e.g. `pub.wide_q1_customer_feedback_8611799`. The id
keeps it unique and gives a stable join back to `pub.surveys` across a rename;
each `pub_layer()` run drops the survey's old view before creating the current
one. Two questions sharing a `shortname` get the question id suffixed onto the
second one's column name.

`survey_wide(con, survey_id)` computes the same shape on demand from `raw`, for
a survey that hasn't been through `pub_layer()` yet.

### Timestamps: verbatim in `raw`, local time in `pub`

`raw` keeps every timestamp exactly as Alchemer sent it — string, suffix and
all. Nothing is parsed or reinterpreted, so the archive never depends on a
judgement of ours being right.

`pub` presents them all as wall-clock time in one timezone, `ALCHEMER_TZ` (or
`pub_layer(tz = )`). Set it to your Alchemer account's timezone, and set it
explicitly for scheduled runs — unset, it falls back to the machine's own
timezone, which would make `pub`'s values depend on which machine last ran.

Alchemer sends two shapes:

- **`date_submitted`/`date_started`** carry a per-row `EST`/`EDT` suffix, so
  they're parsed into an exact instant using that suffix and rendered in
  `ALCHEMER_TZ`.
- **`date_updated`, `created_on`, `modified_on`** carry no suffix and are
  already in the account's own local time, so they're cast as-is. Alchemer's
  documented example makes this plain — one response:

  ```
  date_started   2025-12-09 17:15:30 EST
  date_submitted 2025-12-09 17:15:37 EST
  date_updated   2025-12-09 17:15:43        <- no suffix, same clock
  ```

  Updated 6 seconds after submission. Reading `date_updated` as UTC instead
  would place it 5 hours *before* the response started.

This assumes your account and your analysts share a timezone. If they ever
don't, `raw.responses.date_updated` is still there, verbatim, to convert
however you need.

## `meta`: what the pipeline did

| Table | One row per |
|---|---|
| `meta.runs` | `ingest()` run — status, counts, requests made |
| `meta.run_events` | per-survey decision and refresh, with the reason |
| `meta.survey_state` | survey — change-detection state and failure count |
| `meta.integrity_checks` | assertion result, passes and failures both |
| `meta.loads` | `load_pub_layer()` run, and the tables it wrote |
| `meta.pub_state` | survey built into `pub` — what it was built from, and with |

`db_status()` summarises the latest of these in one row;
`load_pipeline_health()` copies that row into the analytics database for
monitoring. See `vignette("scheduling")`.

## Regional API domains

Set `ALCHEMER_DOMAIN` to match your account's region. Calling the wrong one is,
per Alchemer's documentation, one of the most common causes of failed API calls,
and it fails confusingly — login errors, empty responses, timeouts — rather than
saying so.

| Region | Domain |
|---|---|
| United States | `api.alchemer.com` (default) |
| European Union | `api.alchemer.eu` |
| Canada | `api.alchemer-ca.com` |
| Australia | `api.au.alchemer.com` |
