---
title: "The data model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The data model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## The application database

`ALCHEMER_DB` points at a directory holding two things:

```
$ALCHEMER_DB/
  catalog.ducklake      # the DuckLake catalog -- an ordinary DuckDB file
  data/                  # Parquet data files, written and compacted by DuckDB
```

It's a [DuckLake](https://ducklake.select/) database: Parquet files for
data, plus a small DuckDB file that tracks table structure, snapshots, and
where every row lives. `alchemer_db()` attaches it, under the fixed alias
`alchemer`, as three schemas: `raw`, `meta`, and `pub`.

Only one process may hold the catalog open for writing at a time; any
number of `read_only = TRUE` readers can hold it open simultaneously. If a
write session (`ingest()`, `pub_layer()`, ...) is already running, a *new*
connection attempt -- read or write -- fails immediately with a message
naming the file, rather than hanging; a reader already connected before
that session started is unaffected. See `vignette("scheduling")`'s
Concurrency section for the exact rules, verified against a real database.

## `raw`: archival, verbatim, untyped

Every column in `raw` is `VARCHAR` or `JSON` -- there's no coercion,
renaming, or language selection at ingest time. Alchemer's own typing is
inconsistent enough (`is_test_data` as the string `"1"`/`"0"`, titles as
`{"English": "..."}` maps, the literal statistics key `"total responses"`
with a space) that doing any of that at ingest would let a new question
type or a changed representation fail the whole ingest, or worse, silently
mangle data. Every table also carries a `payload` column holding the
source API object close to verbatim (minus the redactions below), so
nothing this package doesn't model explicitly is ever lost.

| Table | Grain |
|---|---|
| `raw.surveys` | one row per survey |
| `raw.survey_definitions` | one row per survey -- the whole definition tree, verbatim |
| `raw.survey_pages`, `raw.survey_questions`, `raw.survey_question_options` | flattened from that tree |
| `raw.responses` | one row per response; answers verbatim in `survey_data` |
| `raw.survey_statistics`, `raw.survey_campaigns` | one row per question / campaign |

### What's never stored

- **File-upload attachment bytes.** The answer value and Alchemer filename
  are kept, but the bytes themselves are not downloaded, and the pre-signed
  download URL (which expires) is not chased. If you need attachments
  preserved, fetch them yourself while the Alchemer account still exists.
- **Respondent IP address, latitude/longitude, city, region, postal code,
  DMA, and user agent** are redacted at ingest, including inside `payload`.
- **Contacts** are not fetched by `ingest()` (see `?ingest`); they carry
  heavy PII (names, emails, phone numbers) that most uses of this package
  don't need.

## History: DuckLake snapshots, not a hand-rolled archive

There's no separate versioned/archival layer -- DuckLake already versions
every write. Query the database as it stood at a past time or version
directly:

```{r}
con <- alchemer_db()
dplyr::tbl(con, I("alchemer.raw.responses")) # current state

DBI::dbGetQuery(con, "
  SELECT * FROM alchemer.raw.responses AT (TIMESTAMP => '2026-01-01 00:00:00')
")
DBI::dbGetQuery(con, "SELECT * FROM ducklake_snapshots('alchemer')") # every commit
```

**History depth and retention are the same knob.** `expire_history(db,
older_than)` trims how far back time travel can reach (and reclaims the
files that held that history); after that, querying an expired version
fails cleanly rather than returning wrong data. `expunge()` goes further --
it deletes rows and expires their history together, for data that must
actually leave the database, not just become hard to find.

## Deletions are flagged, never dropped

Alchemer's list endpoints silently exclude deleted responses -- there's no
tombstone to detect a deletion from a single request. Instead, every
`ingest()` refresh downloads **all** of a survey's responses and compares
that set to what's already stored: anything previously stored but now
missing from the fetch is flagged `is_deleted = TRUE` and
`deleted_detected_at` is set, never removed. The same happens to a survey
that disappears from account-level discovery. Once flagged, a row stays
flagged unless the API returns it again. Filter deleted rows in or out as
your analysis needs.

## Why full refresh, not incremental

`ingest()` always downloads **every** response for a survey it decides to
refresh -- never a filtered subset. An incremental approach keyed on
`date_updated` was considered and rejected: that field carries no timezone
suffix (Alchemer's own docs call it "typically UTC, but may vary by account
settings"), newly submitted data can take up to five minutes to become
visible via the API, identical requests are cached for 60 seconds, and
deleted responses are silently excluded from list results regardless. Any
one of those can make an incremental delta miss a row permanently, with
nothing downstream able to notice or reconcile it.

Under full refresh, every one of those failure modes degrades only to "this
survey refreshes later than ideal," and self-heals the next time it runs.
`date_updated`, the survey list's `modified_on`, and a per-survey response
count are still used, but only as hints for **when** a survey is worth
refreshing -- they are never used to decide **which rows** to keep. A wrong
hint costs freshness, bounded by `full_sweep_days`; it can't cost
correctness.

## `pub`: typed, for ordinary analysis

`pub_layer()` builds typed, English-resolved (or `language = `-resolved)
tables from `raw` alone -- it's rebuildable, incremental by survey, and
safe to re-run, since it doesn't need archival-grade fidelity the way `raw`
does.

| Table | Grain | Notes |
|---|---|---|
| `pub.surveys`, `pub.questions`, `pub.options` | one row per entity | typed, language-resolved titles |
| `pub.responses` | one row per response | typed timestamps/booleans; test/deleted rows flagged, not dropped |
| `pub.answers` | one row per response x question | question/option labels and reporting values denormalised in |
| `pub.wide_<slug>_<survey_id>` | one row per respondent | a generated view, one per survey |

The wide view's name comes from a lowercased, underscore-separated,
ASCII-transliterated survey title plus the survey id, e.g.
`pub.wide_q1_customer_feedback_8611799` -- the id keeps it unique and gives
a stable join back to `pub.surveys` even after a rename (each
`pub_layer()` run drops a survey's old wide view before creating the
current one). Two questions with the same `shortname` get disambiguated by
suffixing the question id onto the second one's column name.

If you haven't run `pub_layer()` yet for a survey, `survey_wide(con,
survey_id)` computes the same shape directly from `raw` on demand.

## Regional API domains

Set `ALCHEMER_DOMAIN` to match your account's region -- calling the wrong
one is, per Alchemer's documentation, one of the most common causes of
failed API calls, and fails in confusing ways (login errors, empty
responses, timeouts) rather than a clear "wrong domain" message.

| Region | Domain |
|---|---|
| United States | `api.alchemer.com` (default) |
| European Union | `api.alchemer.eu` |
| Canada | `api.alchemer-ca.com` |
| Australia | `api.au.alchemer.com` |
