---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Configure

alchemeR reads everything from the environment, never from arguments in your
project source. Copy
`system.file("extdata", "Renviron.example", package = "alchemeR")` to
`.Renviron` (project) or `~/.Renviron` (user), and fill in your token and
secret from **Account > Integrations > API Key**:

```
ALCHEMER_API_TOKEN=...
ALCHEMER_API_SECRET=...
ALCHEMER_DB=~/alchemer-db
ALCHEMER_TZ=America/Toronto
```

Restart R (or call `readRenviron(".Renviron")`) to pick them up.

`ALCHEMER_DB` is a directory that will hold the archive. `ALCHEMER_TZ` is the
timezone `pub`'s timestamps are presented in — set it to your Alchemer
account's timezone. If your account is outside the US, also set
`ALCHEMER_DOMAIN`; see `vignette("data-model")` for the regional list. Getting
the domain wrong is, per Alchemer's own documentation, one of the most common
causes of failed API calls.

## First run

A cold `ingest()` downloads your **whole** account, which can be a lot of
requests. See what it would do first:

```{r}
library(alchemeR)

ingest(dry_run = TRUE)
```

That makes the same discovery and probe requests as a real run but writes no
data. Then try one small survey:

```{r}
ingest(surveys = "YOUR_SMALL_SURVEY_ID")
db_status()
db_check()
```

`db_status()` summarises what is stored and when it last ran. `db_check()`
re-runs the integrity assertions `ingest()` already ran after that survey
committed — cheap reassurance before trusting the database with everything.

## The real thing

```{r}
ingest()      # every survey in the account
pub_layer()   # tidy typed tables + a wide view per survey
```

Later `ingest()` calls only refresh surveys that actually changed, so it is
safe to run on a schedule — see `vignette("scheduling")`.

## Using the data

`pub` is ordinary tidy data. Use dplyr against it, or pull it into memory:

```{r}
con <- alchemer_db(read_only = TRUE)

# one row per response x question
alchemer_tbl(con, "pub.answers") |>
  dplyr::filter(!is_deleted, !is.na(answer)) |>
  dplyr::collect()

# one row per respondent, the shape most analysts want
survey_wide(con, "YOUR_SURVEY_ID")

DBI::dbDisconnect(con, shutdown = TRUE)
```

Use `read_only = TRUE` whenever you are only reading, and disconnect when you
are done: an open read/write connection stops a scheduled `ingest()` from
running at all, and an open read-only one delays it (see
`vignette("scheduling")`).

`survey_wide()` computes the wide shape on demand, straight from `raw` — handy
for a survey you haven't run `pub_layer()` on yet.
