---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Configure

alchemeR reads everything from the environment, never from arguments in your
project source. Copy
`system.file("extdata", "Renviron.example", package = "alchemeR")` to
`.Renviron` (project) or `~/.Renviron` (user), and fill in your token and
secret from **Account > Integrations > API Key**:

```
ALCHEMER_API_TOKEN=...
ALCHEMER_API_SECRET=...
ALCHEMER_DB=~/alchemer-db
ALCHEMER_TZ=America/Toronto
```

Restart R (or call `readRenviron(".Renviron")`) to pick them up.

`ALCHEMER_DB` is a directory that will hold the archive. `ALCHEMER_TZ` is the
timezone `pub`'s timestamps are presented in — set it to your Alchemer
account's timezone. If your account is outside the US, also set
`ALCHEMER_DOMAIN`; see `vignette("data-model")` for the regional list. Getting
the domain wrong is, per Alchemer's own documentation, one of the most common
causes of failed API calls.

## First run

A cold `ingest()` downloads your **whole** account, which can be a lot of
requests. See what it would do first:

```{r}
library(alchemeR)

ingest(dry_run = TRUE)
```

That makes the same discovery and probe requests as a real run but writes no
data. Then try one small survey:

```{r}
ingest(surveys = "YOUR_SMALL_SURVEY_ID")
db_status()
db_check()
```

`db_status()` summarises what is stored and when it last ran. `db_check()`
re-runs the integrity assertions `ingest()` already ran after that survey
committed — cheap reassurance before trusting the database with everything.

## The real thing

```{r}
ingest()      # every survey in the account
pub_layer()   # tidy typed tables + a wide view per survey
```

Later `ingest()` calls only refresh surveys that actually changed, so it is
safe to run on a schedule — see `vignette("scheduling")`.

## Connecting

alchemeR has no `connect()` helper — you open the connection yourself with
plain `DBI::dbConnect()`, the same as you would for any other database:

```{r}
db <- alchemer_db_path()  # or wherever ALCHEMER_DB points

con <- DBI::dbConnect(duckdb::duckdb())
for (ext in c("json", "sqlite", "ducklake", "icu")) {
  DBI::dbExecute(con, paste("INSTALL", ext))
  DBI::dbExecute(con, paste("LOAD", ext))
}
DBI::dbExecute(con, glue::glue(
  "ATTACH 'ducklake:sqlite:{db}/catalog.sqlite' AS alchemer ",
  "(DATA_PATH '{db}/data/', READ_ONLY)"
))
DBI::dbExecute(con, "USE alchemer")
```

Drop `READ_ONLY` if you need to write (the pipeline functions do this for
you). Typing the real `dbConnect()`/`ATTACH` calls, instead of hiding them
behind a package function, is deliberate: it's what lets RStudio's
Connections pane pick up the connection and let you browse `alchemer`'s
schemas and tables, which it can't do for a connection opened inside
someone else's helper function.

`USE alchemer` makes `alchemer` the default catalog for the rest of the
session, so every query below can say `pub.answers` instead of
`alchemer.pub.answers`. It doesn't collapse `raw`, `meta`, and `pub` into
one search path — `raw` and `pub` deliberately share table names
(`raw.responses` vs `pub.responses`), so the schema always has to be
written out; only the catalog prefix goes away.

You can query the database while a scheduled `ingest()` is running — a
`READ_ONLY` attach never blocks a writer, or is blocked by one.

## Using the data

`pub` is ordinary tidy data. Query it with plain SQL, or with dplyr —
there's no package-specific `tbl()` wrapper, just stock `dbplyr` (install it
alongside alchemeR if you want the pipe; `dplyr::tbl()` only dispatches on a
`DBI` connection once dbplyr's namespace is loaded):

```{r}
# raw SQL
DBI::dbGetQuery(con, "
  SELECT * FROM pub.answers WHERE NOT is_deleted AND answer IS NOT NULL")

# the same thing with dplyr, piped
requireNamespace("dbplyr")
dplyr::tbl(con, dbplyr::in_schema("pub", "answers")) |>
  dplyr::filter(!is_deleted, !is.na(answer)) |>
  dplyr::collect()

# one row per respondent, the shape most analysts want
survey_wide(con, "YOUR_SURVEY_ID")

DBI::dbDisconnect(con, shutdown = TRUE)
```

`survey_wide()` computes the wide shape on demand, straight from `raw` — handy
for a survey you haven't run `pub_layer()` on yet.
