---
title: "Getting started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Configure credentials

alchemeR reads everything from the environment, never from arguments in your
project source. Copy `system.file("extdata", "Renviron.example", package = "alchemeR")`
to `.Renviron` (project) or `~/.Renviron` (user), and fill in your token and secret
from **Account > Integrations > API Key**:

```
ALCHEMER_API_TOKEN=...
ALCHEMER_API_SECRET=...
ALCHEMER_DB=~/alchemer-db
```

Restart R (or call `readRenviron(".Renviron")`) so the variables are picked up.

If your account is outside the US, also set `ALCHEMER_DOMAIN` -- see
`vignette("data-model")` for the regional domain list. Getting this wrong is,
per Alchemer's own documentation, one of the most common causes of failed API
calls.

## First run: a dry run, then one small survey

`ingest()` downloads your **whole** account on a cold run, which can mean a lot
of requests if you have a large or very active account. Before running it for
real, see what it *would* do:

```{r}
library(alchemeR)

ingest(dry_run = TRUE)
```

This makes the same discovery and probe requests as a real run, but writes no
data (it will create the empty database if `ALCHEMER_DB` doesn't exist yet).
Once you're comfortable with the plan, try one small survey first:

```{r}
ingest(surveys = "YOUR_SMALL_SURVEY_ID")
```

Then check what landed:

```{r}
db_status()
db_check()
```

`db_status()` summarises survey/response counts and the most recent run.
`db_check()` runs the integrity assertions `ingest()` already ran after that
survey's commit -- useful confirmation that nothing is broken before trusting
the database with everything.

## The real thing

```{r}
ingest()      # every survey in the account
pub_layer()   # typed tables + a wide view per survey
```

Subsequent calls to `ingest()` only refresh surveys that have actually
changed (or that haven't been refreshed successfully in a while -- see
`vignette("data-model")` for exactly how that's decided), so it's safe to run
on a schedule. See `vignette("scheduling")` for that.

## Using the data

```{r}
con <- alchemer_db()

# Tidy, typed, one row per response x question:
dplyr::collect(alchemer_tbl(con, "pub.answers"))

# One row per respondent, the shape most analysts expect:
survey_wide(con, "YOUR_SURVEY_ID")

DBI::dbDisconnect(con, shutdown = TRUE)
```

If you haven't run `pub_layer()` for a survey yet (or its wide view is
stale), `survey_wide()` computes the same shape directly from `raw` on
demand -- no need to wait for the next `pub_layer()` run.
