---
title: "When an ingest fails"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{When an ingest fails}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Start here

```{r}
library(alchemeR)

ingest_failures()
```

One row per survey whose last refresh failed, worst first, with the reason
attached. Zero rows means nothing is currently failing.

```
# A tibble: 2 x 10
  survey_id title            consecutive_failures last_successful_refresh_at ...
  <chr>     <chr>                           <int> <dttm>
1 7392011   Exit Survey 2025                    9 2026-08-01 03:15:22
2 7418330   Staff Pulse                         9 2026-08-01 03:16:04
```

Read that row by row:

- **`consecutive_failures`** is the shape of the problem. Climbing on every run
  means a *persistent* cause — something about those specific surveys, or about
  the account. A 1 that clears itself next run was a blip.

  A count **stuck at 1 across many runs**, on a survey that is plainly not
  current, means something different: it predates the 1.0.0 fix for failures
  that parked a survey instead of retrying it (see `NEWS.md`). Refresh those
  surveys once explicitly and the count starts telling the truth again:

  ```{r}
  ingest(surveys = c("7392011", "7418330"))
  ```
- **`last_successful_refresh_at`** is how stale the archived data is. It is also
  the reassurance: the data as of that timestamp **is still there and still
  correct**. Each survey refreshes inside one transaction, so a failure rolls
  it back whole — it cannot leave a survey half-updated.
- **`http_status`**, **`message`**, and **`failed_checks`** are the diagnosis.
  The next two sections decode them.

## What a failure has and hasn't done

Worth being clear about before you start digging, because it changes how urgent
this is:

- **The failure is scoped to one survey.** Its refresh runs in its own
  transaction; every other survey in the run still committed.
- **Nothing was lost.** The rollback restores the survey exactly as it was, and
  `raw` never deletes rows anyway — an upstream deletion is flagged, not removed
  (ADR-005).
- **It retries by itself.** A failure does not advance the survey's
  change-detection state, so the next run still sees a difference and refreshes
  it again — and keeps doing so until it succeeds. You do not have to intervene
  to get it retried; you intervene to stop it failing.
- **`pub` still reflects the last good `raw`.** `pub_layer()` rebuilds from
  whatever is archived, so downstream consumers see stale-but-consistent data
  for that survey, never a partial one.

So a handful of failures on one run is noise. The same surveys failing every
run is the thing to chase — and that is what `consecutive_failures` tells you.

## Reading the message

Four shapes turn up, and each points somewhere different.

### `Alchemer API request failed. ... HTTP status: N`

The request itself failed. 429 and the 5xx family are already retried with
exponential backoff — five attempts in total — before the error surfaces, so
seeing one of those statuses here means it failed *persistently*, not once.

| Status | Almost always |
|---|---|
| 401, 403 | Credentials. See "Whole-account failures" below. |
| 404 | The survey vanished between discovery and refresh — usually self-correcting; it will be flagged deleted on the next run. |
| 400 | Frequently the *aftershock* of a timeout — see below. |
| 500 | Alchemer's own 30-second response timeout, if it's the same big survey every time. |
| `unknown (connection failure)` | Network, DNS, or proxy — not Alchemer. |

The message carries the request path with its query string stripped, so you can
see *which* endpoint failed (`/surveyresponse`, `/surveyquestion`, …) without
the credentials ever reaching a log.

### `Alchemer API returned an error. ... Code: N Message: ...`

HTTP said 200; the envelope said `result_ok: false`. `Code` is Alchemer's own,
and it is HTTP-shaped — `401` here is "Login failed / Invalid auth token"
arriving inside a 200 response. `Message` is Alchemer's text verbatim.

### `integrity check(s) failed: ...`

The fetch worked and the write worked, but what landed did not satisfy an
assertion, so the refresh was rolled back rather than committed (ADR-006 —
DuckLake declares no primary keys, so these properties are asserted rather than
enforced). `failed_checks` names them:

- **`no_duplicate_responses`** — the same `response_id` arrived twice. The
  classic cause is *unstable pagination*: `ingest()` pages through
  `/surveyresponse` 500 at a time, and if responses are submitted or edited
  mid-fetch, a row can shift across a page boundary and be read twice.
- **`response_count_matches_fetch`** — the number of live rows stored disagrees
  with the number the refresh believed it fetched. Usually the same cause, seen
  from the other side.
- **`responses_reference_known_surveys`** / **`options_reference_known_questions`**
  — a referential gap, which is unusual and worth reporting.

For the pagination ones, first just **re-run that survey** — the race is
timing-dependent and often does not repeat:

```{r}
ingest(surveys = "7392011")
```

If it repeats every run, the survey is busy enough that a fetch never sees a
quiet moment. Ingesting it at a quieter hour is the practical workaround.

### Anything else

A message that looks like neither an API error nor an assertion is a bug in the
package or a database-level problem. `meta.run_events.message` has the full
text; that is what to include in an issue.

The shape to recognise is a message that names R machinery rather than
Alchemer — `Result must be length 1, not 2`, `Tibble columns must have
compatible sizes`, `subscript out of bounds`. That is a *parse* failure: the
API answered, and the package could not read the answer. Its signature is a
small, fixed set of surveys failing on every run while the rest of the account
is fine, because it is a property of what those surveys contain, not of the
network or the credentials.

Parsing is meant to be incapable of this — `raw` stores every field as text or
JSON precisely so an unfamiliar shape becomes a value rather than an error
(ADR-003). So a parse failure is always a bug worth reporting, and the survey
id plus the message is enough to fix it.

## Whole-account failures

If **every** survey is failing rather than two, stop reading per-survey detail —
the cause is account-wide. In rough order of likelihood:

1. **Wrong regional domain.** Alchemer's own docs call this "one of the most
   common causes of failed API calls", and it fails with misleading symptoms:
   "Login failed / Invalid auth token", empty responses, or connection timeouts,
   all with perfectly valid credentials. Set `ALCHEMER_DOMAIN` to match your
   account — `api.alchemer.com` (US), `api.alchemer.eu`, `api.alchemer-ca.com`,
   `api.au.alchemer.com`.
2. **The key was replaced.** Generating a new API key for a user *immediately
   invalidates the old one*, and keys are per-user — a colleague regenerating
   theirs does not affect yours, but someone regenerating the pipeline's key
   silently breaks it.
3. **API access restricted or disabled.** Account admins can restrict which
   methods are permitted (Account > Integrations > Manage API), and Alchemer
   disables API access outright for repeated rate-limit violations.
4. **Rate limiting.** The account-wide ceiling is 240 requests/minute; the
   default `ALCHEMER_RPM` of 100 deliberately leaves headroom for interactive
   use of the same account. If you have raised it, or several jobs share the
   account, lower it.

A failure *before* any survey is reached — discovery itself — never gets a
per-survey record, because the run never got that far. Its signature is a
`meta.runs` row stuck at `status = 'running'`:

`con` here is a connection you open yourself — see "Connecting" in
`vignette("getting-started")` for the few lines that build one. Every query
below is given as both SQL and stock dbplyr (which you install yourself; it
isn't a dependency of this package), so use whichever you prefer:

```{r}
# SQL
DBI::dbGetQuery(con, "SELECT * FROM meta.runs WHERE status = 'running'")

# dbplyr
dplyr::tbl(con, dbplyr::in_schema("meta", "runs")) |>
  dplyr::filter(status == "running") |>
  dplyr::collect()
```

That means the run died mid-flight — a failed survey-list request, a killed
process, a machine rebooted under a cron job. The console or log output of that
run is the only record of why, which is a good reason for the crontab lines in
`vignette("scheduling")` to redirect output to a file.

## The same surveys, every run

The pattern worth its own section, because it has a specific and non-obvious
cause. If a couple of surveys fail on every run with a **500**, and they are
your **largest** surveys, the likely explanation is Alchemer's **30-second
response timeout**: a page of 500 responses from a survey with very wide or
very numerous questions can take longer than that to generate. Alchemer answers
500, and — per its own documentation — returns **400 for the following minute**,
which is why the retries do not rescue it.

The signature is distinctive: the same survey ids every run, a rising
`consecutive_failures`, a `message` naming `/surveyresponse`, and no problem at
all with the smaller surveys in the same account.

To confirm it, fetch the survey by hand at `ingest()`'s page size and then at a
much smaller one. `alchemer_responses()` passes query parameters through, so
you can vary exactly the thing under suspicion:

```{r}
client <- alchemer_client()

# What ingest() asks for: 500 responses in one request.
system.time(alchemer_responses("7392011", client = client, resultsperpage = 500))

# The same data in smaller bites. Wait a minute in between -- Alchemer caches
# identical GETs for 60 seconds, so back-to-back attempts measure the cache.
system.time(alchemer_responses("7392011", client = client, resultsperpage = 50))
```

A first call that errors or takes anything near 30 seconds while the second
succeeds comfortably confirms the timeout. Compare against a survey that
ingests fine, at the same page size, to rule out something specific to these
two.

If that is the cause, note that `ingest()` itself has no page-size setting
today — it always asks for 500. Until it does, the workable mitigations are to
run those surveys when the account is otherwise quiet, and to raise it with
Alchemer support for the specific survey. Please also open an issue with the
survey's response and question counts: a configurable page size is the right
fix, and that is the evidence for making it.

## Where the detail is kept

`ingest_failures()` answers the standing question. Everything behind it is in
the `meta` schema, and none of it is inside the transaction that gets rolled
back (ADR-007) — failure records are written precisely so they outlive the
failure:

| Table | One row per | Use it for |
|---|---|---|
| `meta.runs` | `ingest()` run | Did the run finish? How many failed? How many requests did it cost? |
| `meta.run_events` | decision, probe, refresh, run | The full log — including probes that failed without failing their survey |
| `meta.survey_state` | survey | Current per-survey state: `consecutive_failures`, last success |
| `meta.integrity_checks` | assertion | Every check, **pass and fail**, from every run |
| `meta.loads` | `load_pub_layer()` run | The Load stage, if you use it |

Query them like any other table:

```{r}
# --- SQL ---

# Everything that went wrong in the last 24 hours, newest first
DBI::dbGetQuery(con, "
  SELECT * FROM meta.run_events
  WHERE status = 'error' AND finished_at > now() - INTERVAL 1 DAY
  ORDER BY finished_at DESC")

# One survey's whole history, decisions included: is it failing, or was it
# never being refreshed in the first place?
DBI::dbGetQuery(con, "
  SELECT * FROM meta.run_events
  WHERE survey_id = '7392011' ORDER BY finished_at DESC")

# Probe failures that did not fail their survey -- they cost requests and
# defeat change detection, so a survey with many is being refreshed on every
# single run for no reason.
DBI::dbGetQuery(con, "
  SELECT survey_id, count(*) AS n FROM meta.run_events
  WHERE phase = 'probe' AND status = 'error'
  GROUP BY survey_id ORDER BY n DESC")
```

The same three, through dbplyr:

```{r}
events <- dplyr::tbl(con, dbplyr::in_schema("meta", "run_events"))

events |>
  dplyr::filter(status == "error", finished_at > Sys.time() - 86400) |>
  dplyr::arrange(dplyr::desc(finished_at)) |>
  dplyr::collect()

events |>
  dplyr::filter(survey_id == "7392011") |>
  dplyr::arrange(dplyr::desc(finished_at)) |>
  dplyr::collect()

events |>
  dplyr::filter(phase == "probe", status == "error") |>
  dplyr::count(survey_id) |>
  dplyr::collect()
```

`meta.run_events` grows by a few rows per survey per run and is never trimmed by
`compact()` or `expire_history()`, so the history is there for as long as the
database is.

## Investigating one survey

```{r}
# 1. Reproduce it, on that survey alone. `surveys=` always refreshes,
#    regardless of change detection, and the reason comes back in `message`.
out <- ingest(surveys = "7392011")
out$message

# 2. Is it the refresh, or was the data already inconsistent? These are the
#    same assertions ingest() runs, account-wide and read-only.
db_check()

# 3. What is actually archived for it right now? Either of these:
DBI::dbGetQuery(con, "
  SELECT is_deleted, count(*) AS n FROM raw.responses
  WHERE survey_id = '7392011' GROUP BY is_deleted")

dplyr::tbl(con, dbplyr::in_schema("raw", "responses")) |>
  dplyr::filter(survey_id == "7392011") |>
  dplyr::count(is_deleted) |>
  dplyr::collect()
```

`ingest(dry_run = TRUE)` is the safer first step when you are not sure which
surveys are involved: it makes the same discovery and probe requests and reports
the same decisions, but writes nothing.

If a refresh has committed something you want to inspect against its
predecessor, the archive is versioned — `vignette("data-model")` covers time
travel, which lets you query the survey as of the snapshot before the change.

## Getting alerted instead of noticing

`ingest()` now warns when any survey failed, so a scheduled run's log says so
without anyone reading a tibble. Beyond that:

- `inst/scripts/ingest_job.R` and `etl_pipeline.R` both print the failing
  surveys with their messages — use them as the basis for your own job.
- `db_status()` is the one-row summary; `last_run_status` is
  `"completed_with_errors"` when any survey failed.
- `load_pipeline_health()` copies that row into your analytics database, so an
  alert on "status is not completed" or "last load older than X" needs no access
  back to `ALCHEMER_DB`. See `vignette("scheduling")`.
